---
title: "Similarity, uniqueness of Swedish radio stations"
author: "Martin Rydén"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(dplyr)

```
Last updated on: `r Sys.time()`

See **radio.R** for full script including data scraping. Presented data is from past 7 days (July 2nd). Song identity is assumed based on name - different songs with same name will be counted as the same. Scraped data may be incorrect due to spelling variants, missing data, incomplete sources etc. Only publicly available data is used.

```{r, echo = F}
combined <- fread("data/combined.csv")
```

## How many songs were played?
#### Total number of songs (including duplicates) played between 2021-06-26 and 2021-07-02

```{r, echo = F}
count_by_station <- combined[, .N, by = "Station"]
names(count_by_station) <- c("Station", "Number of songs")
count_by_station[order(`Number of songs`, decreasing = T)] %>% knitr::kable()
```

## Uniqueness
#### Percentage of unique songs (within stations) -- low % means high repetitiveness

```{r, echo = F}
unique_by_station <- unique(combined, by = c("Station", "Title"))

unique_by_station <- unique_by_station[, .N, by = "Station"]

unique_by_station$Total <- count_by_station$N

unique_by_station$Ratio <- round(unique_by_station$N/unique_by_station$Total * 100)

names(unique_by_station) <- c("Station", "Unique songs", "All songs", "% unique")

unique_by_station[order(`% unique`, decreasing = T)] %>% knitr::kable()
```

## Similarity
#### What radio stations played the same songs the most?

```{r, echo = F, fig.width=8, fig.height=9, fig.cap = "Counts of same song being played between sets of stations, highest degree of similarity between P3 and P4."}

# Group songs by Station
songs_by_group <- combined[,list(list(unique(Station))), by="Title"]

# Which songs have been played by more than one station?
songs_by_group <- songs_by_group[lengths(songs_by_group$V1) > 1]

dt = rbindlist(
	lapply(songs_by_group$V1, function(x) data.table(t(x))),
	fill = TRUE
)

combs <- t(combn(sort(na.omit(unique(unlist(dt)))), 2))
freqs <- apply(combs, 1, function(C) {
	sum(apply(dt, 1, function(a) all(C %in% a, na.rm = TRUE)))
})
combsDF <- as.data.frame(combs)
combsDF$freq <- freqs

combsnew <- data.table(paste0(combsDF$V1, " - ", combsDF$V2), combsDF$freq)

combsnew$setby <- combsDF$V1

library(ggplot2)
library(ggthemes)
library(RColorBrewer)

ggplot(combsnew, aes(x = V1, y = V2, fill = setby)) +
	scale_fill_manual(values = brewer.pal(n = 11, name = "Set3"), name = "") +
	xlab("Stations") + ylab("# times played same song (i.e. similarity)") +
	coord_flip() +
	geom_bar(stat = "identity") +
	ggthemes::theme_few(base_size = 14)
	

```



## Similarity heatmap

```{r, echo = F, message = F, warnings = F, error = F, results = "hide", message = "F", fig.show="hold", out.width="50%", fig.cap = "Similarity heatmap based on counts of same song played between sets of radio stations. Log-transformed and centered plots to emphasize dissimilarity."}

library(pheatmap)

M <- reshape2::dcast(combsDF, V2~V1, value.var="freq", drop = F)

rownames(M) <- M$V2
M <- M[, 2:ncol(M)]

M <- M[ , order(names(M))]
M <- M[order(rownames(M)),]

pheatmap(M, cluster_rows=F, cluster_cols=F, na_col="white", main = "Raw counts")
pheatmap(log10(M+1), cluster_rows=F, cluster_cols=F, na_col="white", main = "log10 counts")
pheatmap(scale(M, center = T, scale = F), cluster_rows=F, cluster_cols=F, na_col="white", main = "Centered counts")


```



## Timepoint analysis
#### Are certain songs played at a certain time?

For each group of stations, we investigate their shared top 12 songs and plot them as ridge plots (density/24h).

```{r, echo = F, fig.show="hold", out.width="50%"}
combined$time <- as.POSIXct(strptime(combined$Timestamp, format = "%H:%M")) 

library(ggridges)
library(scales)

lims <- as.POSIXct(strptime(c("00:00", "23:59"), 
							format = "%H:%M"))

joy_plot <- function(stations) {
	combined_top <- combined[Station %in% stations]
	combined_top_titles <- combined_top[, .N, by = "Title"]
	combined_top_titles <- combined_top_titles[order(N, decreasing = T)][1:12]
	combined_top <- combined_top[Title %in% combined_top_titles$Title]
	combined_top$Title <- as.factor(combined_top$Title)
	combined_top$Station <- as.factor(combined_top$Station)
	
	ggplot(combined_top, aes(x = time, y = Title, fill = Title, height = ..scaled..)) +
		geom_density_ridges(stat = "density", trim = F, scale = 0.6) +
		scale_x_datetime(breaks = date_breaks("4 hours"), labels = date_format("%H:%M"), limits = lims) +
		scale_fill_manual(values = brewer.pal(n = 12, name = "Set3"), name = "") +
		
		guides(fill = F) +
		theme_clean() +
		theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
		facet_grid(~ Station)
}

```

#### Pop, miscellaneous
```{r, echo = F, fig.show="hold", out.width="50%"}

joy_plot(c("P3", "P4", "NRJ", "Rix FM", "Mix Megapol"))
joy_plot(c("P3", "Din Gata"))
joy_plot(c("Lugna Favoriter", "NRJ", "P4", "Mix Megapol"))

```

#### Classical

```{r, echo = F}

joy_plot("P2")

```

#### Rock

```{r, echo = F}
joy_plot(c("Rockklassiker", "Bandit Rock", "Retro FM", "Pirate Rock"))

```


```{r, echo = F, fig.show="hold", out.width="50%"}
combined$time <- as.POSIXct(strptime(combined$Timestamp, format = "%H:%M")) 

library(ggridges)
library(scales)

lims <- as.POSIXct(strptime(c("00:00", "23:59"), 
							format = "%H:%M"))

joy_plot2 <- function(station1, station2) {
	# For both stations, find most popular songs
	combined_top1 <- combined[Station == station1, .N, by = "Title"]
	combined_top2 <- combined[Station == station2, .N, by = "Title"]
	
	# We arbitrarily decide to keep songs played more than three times
	combined_top1 <- combined_top1[N > 2]
	combined_top2 <- combined_top2[N > 2]
	
	combined_top1 <- combined_top1[order(N, decreasing = T)]
	
	# Check every title from station 1 to see if it's also among popular songs in station 2
	combined_top_shared <- combined_top1[Title %in% combined_top2$Title]
	
	combined_top_shared <- combined[(Title %in% combined_top_shared$Title) & (Station %in% c(station1, station2))]
	
	combined_top_shared <- combined_top_shared[, .N, by = "Title"]
	combined_top_shared <- combined_top_shared[order(N, decreasing = T)][1:12]
	
	combined_top_shared <- combined_top_shared[Title %in% combined_top_shared$Title]
	
	combined_top_shared <- combined[(Title %in% combined_top_shared$Title) & (Station %in% c(station1, station2))]
	
	ggplot(combined_top_shared, aes(x = time, y = Title, fill = Title, height = ..scaled..)) +
		geom_density_ridges(stat = "density", trim = F, scale = 0.6) +
		scale_x_datetime(breaks = date_breaks("4 hours"), labels = date_format("%H:%M"), limits = lims) +
		scale_fill_manual(values = brewer.pal(n = 12, name = "Set3"), name = "") +
		
		guides(fill = F) +
		theme_clean() +
		theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
		facet_grid(~ Station)
}

```

Above plots show shared top 12 songs but has a bias toward stations that play more songs and is more repetitive. We can reduce the plots to only show songs that has been played at least a few times on each station.

#### Pairwise comparisons, shared titles (inclusive)

```{r, echo = F, out.width="50%"}
joy_plot2("Retro FM", "Lugna Favoriter")
joy_plot2(station1 = "P3", station2 =  "P4")
joy_plot2(station1 = "P2", station2 =  "P4")
joy_plot2(station1 = "P3", station2 =  "Din Gata")
joy_plot2(station1 = "Rockklassiker", station2 =  "Bandit Rock")
joy_plot2(station1 = "Rockklassiker", station2 =  "Retro FM")



```